Palomar 5 Gaps

### 2024-10-30 paper/main.tex
added comments post disucssion with Paola

  As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales. A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos. For instance, \citet{2008ApJ...679.1260M} computed cosmological simulations and compared the Via Lactea simulations \citep{2008ApJ...679.1260M} and demonstrated (\salvatore{I don't quite understand Madau 2008. The main result is that based on the assumptions you can produce a variance in the number of subhaloes that are consistent with global properties of a milky-way like galaxy? I need to discuss this with another human}). \citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modelign such physical phenomena. \citet{2019Galax...7...81Z} rightly states that until particles are discovered $\Lambda$-CDM remains a hypothesis. Since the true nature of dark matter is elusive. The possibilities for sub-galactic behaviour and structure of dark matter are vaste. However, as proposed by Carlberg in 2012, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.

@@

  Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and spectroscopy. The development and deployment of the Streamfinder algorithm has led to the detection of about 90 stellar streams in the Milky Way. \citet{2023MNRAS.520.5225M} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters. Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres. The solo-streams can be explained by a larger globular cluster population whose memebers have been disolved due to two body relaxation and tidal stripping. For instance, Z Wan 2020(???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise. This implies that the stream is from a completely disolved globular cluster which comes from a population of clusters that have not survived until modern times.

@@

@@

>>>>

  As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales. A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos. For instance, \citet{2008ApJ...679.1260M} computed cosmological simulations and compared the Via Lactea simulations \citep{2008ApJ...679.1260M} and demonstrated (\salvatore{I don't quite understand Madau 2008. The main result is that based on the assumptions you can produce a variance in the number of subhaloes that are consistent with global properties of a milky-way like galaxy? I need to discuss this with another human}). \citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modeling such physical phenomena. \salvatore{OK. The most important point from Zavala to highlight is there information baked within a population of dark matter haloes, cioè fundamental parameters. Only certain fundamental parameters of dark matter could give a certain population. Having information thus on the population of streams will give information on the nature of dark matter.} \citet{2019Galax...7...81Z} rightly states that until particles are discovered $\Lambda$-CDM remains a hypothesis. Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vaste. However, as proposed by Carlberg in 2012, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.

@@

  Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and spectroscopy. The development and deployment of the Streamfinder algorithm has led to the detection of about 90 stellar streams in the Milky Way. \citet{2023MNRAS.520.5225M} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters. Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres. The solo-streams can be explained by a larger globular cluster population whose memebers have been disolved due to two body relaxation and tidal stripping. For instance, Z Wan 2020( can't find the citation???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise. This implies that the stream is from a completely disolved globular cluster which comes from a population of clusters that have not survived until modern times.

@@

    \salvatore{What's the point? In the end were more interested in the streams for probing the galaxy. Having a more accurate model of the internal dynamics Palomar 5 would change the luminosity per square degree. }

@@

  \salvatore{Model limits? lets discuss them here.}
